%----------------------------------------------------------------------------------------
%	CONCLUSION & RECOMMENDATIONS.
%----------------------------------------------------------------------------------------

\section{\label{chap:conclusion}Conclusion \& Recommendations}

The era of the origin of Earth has plenty of mysteries and questions remaining to be answered. One of the most important of these questions is about the origin of water on this planet. The answer to this question would reveal how life started to exist and on how other terrestrial planets in other solar systems might develop. This question of the origin of water has been asked and tried to have a satisfactory answer for several decades, no answer has been completely satisfactory. The main two hypotheses are the Late Heavy Bombardment and the early accretion model. \\

The Late Heavy Bombardment says that Earth was formed dry and received water and volatiles from extraplanetary sources. This model dismisses all claims from the early accretion model concerning the possibility of the Earth forming wet. The early accretion model does not preclude the Late Heavy Bombardment, however does severely limit the amount of water accumulated from extraplanetary sources. 
The issues with Late Heavy Bombardment is that the isotopes of comets do not match Earth's deuterium to hydrogen ratio; the asteroids have issues with the quantity of hydrated materials. The Early Accretion model has issues with retention of water in during the development process.
There is agreement on “our knowledge of detail of terrestrial planet formation and hydration is currently insufficient” \cite{BOMB14}. \\

There has been research and modeling that shows only physical absorption or chemical absorption. There should be new research focus on the combination of those two absorption methods to model how much water would be absorbed into grains and minerals. 
The model should determine whether the grains and hydrated minerals would retain water at elevated temperatures that are seen during the development of terrestrial planets.
There should be further research into seeing if the minerals and dust can retain their water at the pressure associated with the impact nature of accretion.
With more research there can be a possible evidence or discrepancy that could make one scenario more or less likely to have occurred.
