%----------------------------------------------------------------------------------------
%	INTRODUCTION.
%----------------------------------------------------------------------------------------

\section{Introduction}
Scientists aren't sure where the water on Earth and other terrestrial planets originated from. There are multiple explanations of where the water came from. But there is currently no perfect explanation how the terrestrial planets gained water. It is believed that the answer can be found in studying the D/H ratio in comets and asteroids. Because this ratio is directly linked to the place in the solar system the objects formed.\\

The most believed explanation is that the water was imported from comets and/ or asteroids that are rich in water. The proto-Earth would have been quite hot because of the collisions with other protoplanets. In this case the proto-Earth accreted dry. However, it is also possible that the Earth accreted wet, because there was water in the early solar system. In this case there should be a way for the proto-Earth to hold on to this water.\\

In this report the main research question is "Are comets the source of water?". To answer this question firstly the formation of comets and asteroids in the early solar system is being researched in chapter \ref{chap:formation}. In this chapter the measured data from the comet 67P/Churyumov-Gerasimenko is being researched. This is the only comet where a lander did direct measurements. Because there is the only comet, it can't be ruled out that other comets have the same D/H ratio. This is why also measured data from other comets are used. These measurements have higher tolerances because there hasn't been any direct measurements.\\

In chapter \ref{chap:creation} the possibility of the proto-Earth that accreted wet is being researched. The most important question is if the proto-planets could hold on to this water. It is believed that that because of the huge collisions the terrestial planets with other protoplanets there was a lot of heat. But there are theories that the protoplanets could hold on to those volatile elements inside the planet. \\

In the last chapter \ref{chap:bombardment} the options that comets and asteroids brought the water to proto-Earth during the late bombardment. During this time there weren't many large collisions with proto-planets. The proto-Earth could cool down and there were only small impacts. Because of this the atmosphere wouldn't be completely destroyed and could absorb the water vapor. In chapter \ref{chap:conclusion} the conclusion of this research is stated and are recommendations given to get more information about the source of water on Earth.\\ 


